plyranges: a grammar of graphics for genomics data

### 2017-11-22 paper.Rmd
plos format added

---
title: Outline of plyranges a grammar of genomic data manipulation
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
bibliography: references.bib
link-citations: true

---

Overview of structure, references and layout for the plyranges package paper.

# Abstract

The Bioconductor project has created many powerful abstractions for
reasoning about genomics data, such as the _Ranges_ data structures 
for representing genomic intervals. By recognising that these data structures
follow 'tidy' data principles we have created a grammar of genomic data 
manipulation that defines verbs for performing actions on and between genomic 
interval data. This grammar simplifies performing common genomic data analysis 
tasks via method chaining, type consistency and results in creating human 
readable pipelines. We have implemented this grammar as an Bioconductor/R 
package  plyranges.

# Introduction

Putting the work into context.

## What do we mean by genomic data?

@@

Key references:

# Design and Implementation

Discussion of the API and its key features

- every action on a Range is a verb (functional composition, method chaining)
- fluent (human readable, code describes what to do rather than 'how')
- every function returns a class familiar to the user
- an expressive algebra for performing arithmetic (anchoring functions)
- split-apply-combine strategies (group_by + reduce/summarise/mutate/filter)
- recasting overlapping/nearest etc as joins

# Results

A case study showing plyranges makes life easier...
Perhaps could use long reads data from Matt and his post-doc Charity (interesting
use case of looking for intron-exon junctions)

# Conclusion (future directions)

- availability and maintenance of package
- extending to SummarisedExperiment class
- natural fit with ggbio2

# References

>>>>

---
title: "plyranges: a grammar of graphics for genomics data" 
author:
  - name: Stuart Lee
    email: stuart.lee1@monash.edu
    affiliation: Monash University
  - name: Michael Lawrence
    email: michael.lawrence@gene.com
    affiliation: Genentech
  - name: Di Cook
    email: di.cook@monash.edu
    affiliation: Monash University
address:
  - code: Monash University
    address: Department of Econometrics and Business Statistics, Clayton, Victoria, Australia
  - code: Genentech
    address: Bioinformatics and Computational Biology, Genentech, Inc., South San Francisco, California, United States of America 
abstract: |
    The Bioconductor project has created many powerful abstractions for reasoning about genomics data, such as the _Ranges_ data structures for representing genomic intervals. By recognising that these data structures follow 'tidy' data principles we have created a grammar of genomic data manipulation that defines verbs for performing actions on and between genomic interval data. This grammar simplifies performing common genomic data analysis tasks via method chaining, type consistency and results in creating human readable pipelines. We have implemented this grammar as an Bioconductor/R package called plyranges.

bibliography: references.bib
output: rticles::plos_article
csl: plos.csl
---

# Introduction
## What do we mean by genomic data?

@@

# References {#references .unnumbered}
